Robust Spectroscopic Inference with Imperfect Models

### 2014-10-31 tex/paper.tex
new hdf5 format, rerunning gl, wasp

\acknowledgments  The authors would like to graciously acknowledge the following people for many helpful discussions and key insights: Gregory M. Green, Kaisey Mandel, Daniel Foreman-Mackey, John Johnson and the ExoLab, Daniel Eisenstein, Allyson Bieryla, and Maxwell Moe.  IC was supported by the NSF Graduate Fellowship and the Smithsonian Institution. Figures \ref{fig:Kurucz_posterior}, \ref{fig:PHOENIX_posterior}, and \ref{fig:Gl51_posterior} were generated with \texttt{triangle.py} \citep{foreman-mackey14}. Astropy. JuliaLang. Python/Scikit-learn.\\

@@

Interpolation from a library of synthetic spectra is conceptually straightforward, but in the 
application discussed here there are some subtleties that require detailed consideration.  
Spectral line formation is a non-linear process, where line shapes are determined by the complex 
radiative transfer through a stellar atmosphere.  We aim to bypass this non-linear 
process and predict (through interpolation) the behavior of spectral lines based on pre-computed 
library spectra with similar physical parameters.  The goal in this Appendix is to describe a 
procedure by which we propagate an empirical approximation of the {\it uncertainty} in that 
prediction through the probabilistic framework we have developed.

There are two important factors to consider when interpolating in libraries of model stellar 
spectra.  First, the flux values in some non-negligible fraction of the spectra do not vary 
monotonically with changes in the physical parameters that define the library: this helps inform 
the decision about which functional form of interpolation should be used.  And second, the spectra 
are usually not Nyquist sampled with respect to the physical parameters that define the library.  
This unfortunately (but not surprisingly) means that the library is not sampled densely enough in 
parameter-space to reconstruct a synthetic spectrum for arbitrary parameters without some 
information loss.  Given these factors, we choose to develop a spectral ``emulator,'' which, given the information contained in the library of synthesized spectra, seeks to non-deterministically predict the behavior of a spectrum for stellar parameters not sampled by the library.

@@

The first step in designing an emulator is to decompose the spectral library into an appropriate basis for emulation. We use principal component analysis (PCA) to decompose the synthetic library into a set of ``eigenspectra'' $\varphi$. Following techniques of SLOAN (cite Ivecik 2014). After examining how well the spectra reconstruct the grid, we keep the first $m$ of the eigenspectra, and discard the latter eigenspectra. We keep enough eigenspectra to reconstruct any pixel flux to better than $2\%$, however most pixels are more like $XX\%$. An $N_{\rm pix}$ eigenspectrum $\xi_k$, $k = \{1, 2, \ldots, m\}$. Before we do the PCA operation, however, we first standardize the synthetic library by subtracting the mean spectrum and ``whitening'' the spectra by dividing by the standard deviation in each pixel. First, we select only a subset of the spectra synthetic library which will be relevant to our  The mean spectrum is 
\begin{equation}
  \xi_\mu = \frac{1}{N} \sum_{i = 1}^N f_\lambda(\{\vt_\ast \}^\textrm{grid})
\end{equation}
and the standard deviation spectrum is
\begin{equation}
  \xi_\sigma = \sqrt{\frac{1}{N} \sum_{i=1}^N \bigl [ f_\lambda(\{\vt_\ast \}^\textrm{grid}) - \xi_\mu \bigr]^2 } 
\end{equation}
We standardize all spectra in the library by subtracting the mean spectrum and dividing by the standard deviation spectrum.
\begin{equation}
  \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) = \frac{f_\lambda(\{\vt_\ast \}^\textrm{grid}) - \xi_\mu}{\xi_\sigma}
\end{equation}
Then, the eigenspectra are determined from the standardize grid using PCA analysis (see Ivecik, scikit-learn). To speed computation, we first select $M$ subset spectra from the grid.

>>>>

\acknowledgments  The authors would like to graciously acknowledge the following people for many helpful discussions and key insights: Gregory M. Green, Kaisey Mandel, Daniel Foreman-Mackey, John Johnson and the ExoLab, Daniel Eisenstein, Rebecca Dawson, Tom Loredo and the ExoStat group, Allyson Bieryla, and Maxwell Moe.  IC was supported by the NSF Graduate Fellowship and the Smithsonian Institution. Figures \ref{fig:Kurucz_posterior}, \ref{fig:PHOENIX_posterior}, and \ref{fig:Gl51_posterior} were generated with \texttt{triangle.py} \citep{foreman-mackey14}. This research made use of Astropy \citep{astropy13} and the Julia programming language \citep{julia12}.

@@

Interpolation from a library of synthetic spectra is conceptually straightforward, but in the 
application discussed here there are some subtleties that require detailed consideration.  
Spectral line formation is a non-linear process, where line shapes are determined by the complex 
radiative transfer through a stellar atmosphere.  Any interpolation process aims to bypass this non-linear 
process and predict the behavior of spectral lines based on pre-computed 
library spectra with similar physical parameters.  This appendix describes the 
procedure by which we propagate an empirical approximation of the {\it uncertainty} in that 
prediction--which can be substantial--through the probabilistic framework we have developed.

There are two important factors to consider when interpolating in libraries of model stellar 
spectra.  First, the flux values in some non-negligible fraction of the spectral pixels do not vary 
monotonically with changes in the physical parameters that define the library: this helps inform 
the decision about which functional form of interpolation should be used.  And second, the spectra 
are usually not Nyquist sampled with respect to the physical parameters that define the library.  
This unfortunate fact means that the library is not sampled densely enough in 
parameter-space to reconstruct a synthetic spectrum for arbitrary parameters without some 
information loss.  Given these factors, we choose to develop a spectral ``emulator,'' which, given the information contained in the library of synthesized spectra and a chosen interpolation point, delivers a probability distribution over the range of possible interpolated spectra.

@@

The first step in designing an emulator is to decompose the spectral library into an appropriate basis for emulation \citep{habib07, heitmann09}. We chose the principal component basis (PCA) to decompose the synthetic library into a set of ``eigenspectra,'' following the techniques of \citet{ivezic13}. We decide to truncate our basis to the first $m$ of the eigenspectra, where $m$ is decided by the minimum number of eigenspectra required to reproduce any spectrum in the grid to better than 2\% accuracy for all pixels, however the mean pixel error is generally much smaller ($\lesssim 0.5\%$). An $N_{\rm pix}$ eigenspectrum $\xi_k$, $k = \{1, 2, \ldots, m\}$. Before we do the PCA operation, however, we first standardize the synthetic library by subtracting the mean spectrum and ``whitening'' the spectra by dividing by the standard deviation in each pixel. First, we select only a subset of the spectra synthetic library which will be relevant to our  The mean spectrum is

\begin{figure*}[!t]
  \begin{center}
  \includegraphics{figs/pca_reconstruct.pdf}
  \figcaption{A bunch of reconstrcuted.
\label{fig:pca_reconstruct}}
  \end{center}
\end{figure*}

\begin{equation}
  \xi_\mu = \frac{1}{N} \sum_{i = 1}^N f_\lambda(\{\vt_\ast \}^\textrm{grid})
\end{equation}
and the standard deviation spectrum is
\begin{equation}
  \xi_\sigma = \sqrt{\frac{1}{N} \sum_{i=1}^N \bigl [ f_\lambda(\{\vt_\ast \}^\textrm{grid}) - \xi_\mu \bigr]^2 } 
\end{equation}
We standardize all spectra in the library by subtracting the mean spectrum and dividing by the standard deviation spectrum.
\begin{equation}
  \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) = \frac{f_\lambda(\{\vt_\ast \}^\textrm{grid}) - \xi_\mu}{\xi_\sigma}
\end{equation}
Then, the eigenspectra are determined from the standardize grid using PCA analysis (see Ivecik, scikit-learn). To speed computation, we first select $M$ subset spectra from the grid.
